Cooperative Kernels: Safe Multitasking for Blocking Algorithms on GPUs

### 2016-11-14 write-ups/PLDI/paper.tex

\myfiglong\ref{fig:workstealing-cooperative} shows a cooperative
version of the work stealing example of \myfig\ref{fig:workstealing} from \mysec\ref{sec:openclexamples}.
%
In this example there is no state to transmit since a computation is
entirely parameterised by a task, which is retrieved from a queue
located in global memory. In the main loop, before trying to obtain a
task, we make the workgroup offer itself to be forked or killed
(lines~\ref{line:cwksteal:offerfork}-\ref{line:cwksteal:offerkill}). Note
that a workgroup may be killed even if its associated task queue is
not empty, since the remaining tasks will be stolen by other
workgroups.  Care must be taken on computing queue identifiers: the
original version in \myfig\ref{fig:workstealing} can rely on the fact
that there are as many queues as workgroups, but here the number of
workgroups may change during execution. Therefore, the number of
queues must now be explicitely given as an argument
(line~\ref{line:cwksteal:argnbqueue}) and used as an upper bound of
queue indexes when scanning queues in $\mathsf{more\_work}$
(line~\ref{line:cwksteal:morework}) and $\mathsf{pop\_or\_steal}$
(line~\ref{line:cwksteal:poporsteal}). Besides, computation of the
workgroup's queue index (line~\ref{line:cwksteal:getgroupid}) must now
be moved inside the main loop since a forked workgroup that starts to
execute after $\offerfork$ must re-compute its group id.

@@

\begin{lstlisting}
kernel work_stealing(global Task * queues,
(*@\label{line:cwksteal:argnbqueue}@*)                     const int nb_queues) {
(*@\label{line:cwksteal:morework}@*)  while (more_work(queues, nb_queues)) {
(*@\label{line:cwksteal:offerfork}@*)    offer_fork();
(*@\label{line:cwksteal:offerkill}@*)    offer_kill();
(*@\label{line:cwksteal:getgroupid}@*)    int queue_id = get_group_id();
(*@\label{line:cwksteal:poporsteal}@*)    Task * t = pop_or_steal(queues, queue_id, nb_queues);
    if (t) {
      // task processing may create more work
(*@\label{line:cwksteal:processtask}@*)      process_task(t, queues, queue_id);
    }
  }
}
\end{lstlisting}

\caption{Cooperative version of the work stealing kernel of
  \myfig~\ref{fig:workstealing}, using $\offerfork$ and $\offerkill$,
  and passing the total number of queues
  explicitely}\label{fig:workstealing-cooperative}
\end{figure}

>>>>

\myfiglong\ref{fig:workstealing-cooperative} shows a cooperative
version of the work stealing example of \myfig\ref{fig:workstealing} from \mysec\ref{sec:openclexamples}.
%
In this example there is no state to transmit since a computation is
entirely parameterised by a task, which is retrieved from a queue
located in global memory. In the main loop, before trying to obtain a
task, we make the workgroup offer itself to be forked or killed
(lines~\ref{line:cwksteal:offerfork}-\ref{line:cwksteal:offerkill}). Note
that a workgroup may be killed even if its associated task queue is
not empty, since remaining tasks will be stolen by other
workgroups. Since $\offerfork$ may be the entry point of a workgroup,
the queue id must now be computed after it (line~\ref{line:cwksteal:getgroupid}). In particular, the queue
id cannot be transmitted since we want a newly spawned workgroup to
read its own queue and not the one of the forking workgroup.

% The queue id cannot be transmitted since it corresponds to the
% workgroup id.
>>>>>>> 65caf661ed1c47ada958a674a0605c42b495e409

@@

\begin{lstlisting}
kernel work_stealing(global Task * queues) {
(*@\label{line:cwksteal:morework}@*)  while (more_work(queues)) {
(*@\label{line:cwksteal:offerfork}@*)    offer_fork();
(*@\label{line:cwksteal:offerkill}@*)    offer_kill();
(*@\label{line:cwksteal:getgroupid}@*)    int queue_id = get_group_id();
(*@\label{line:cwksteal:poporsteal}@*)    Task * t = pop_or_steal(queues, queue_id);
    if (t) {
      // task processing may create more work
(*@\label{line:cwksteal:processtask}@*)      process_task(t, queues, queue_id);
    }
  }
}
\end{lstlisting}

\caption{Cooperative version of the work stealing kernel of
  \myfig~\ref{fig:workstealing}, using $\offerfork$ and
  $\offerkill$}\label{fig:workstealing-cooperative}
\end{figure}
